---
title: "Comparing Persistence barcodes with the same topology"
author: "Vincent Runge"
date: "05/11/2023"
output:
  html_document:
    keep_md: true
    css: styles.css
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
---


```{r setup, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, message=FALSE}
library(TDA)
library(tdaTS) #our package
library(tibble)
library(plotly) #3D plotting
library(fields) #image.plot function
```


# Objective


We propose to compare the persistence diagram or the barcodes of data generated with the same topology.

We need to find a distance between barcodes close to zero even when the barcodes look different (after some kind a rescaling...)

# Detecting a hole (H1)

```{r}
myplots <- function(data, complex = "alpha")
{
  if(complex == "alpha")
  {
      DiagAlphaCmplx <- alphaComplexDiag(X = data,
                                     library = c("GUDHI", "Dionysus"),
                                     location = TRUE)
  }
  if(complex == "rips")
  {
      DiagAlphaCmplx <- ripsDiag(X = data,
                                maxdimension = 1,
                                maxscale = 1,
                                dist = "euclidean",
                                library = c("GUDHI", "Dionysus"),
                                location = TRUE)
  }

  plot(DiagAlphaCmplx$diagram, barcode = TRUE)
  plot(data, col = 1,xaxt="n", yaxt="n",xlab="", ylab="", asp = 1)
  one <- which(DiagAlphaCmplx[["diagram"]][, 1] == 1)
  for (i in seq(along = one))
  {
  for (j in seq_len(dim(DiagAlphaCmplx[["cycleLocation"]][[one[i]]])[1]))
    {lines(DiagAlphaCmplx[["cycleLocation"]][[one[i]]][j, , ] + rnorm(n = 4,sd = 0.0),cex = 1, col = i + 1)}
  }
  return(DiagAlphaCmplx$diagram)
}
```



## Square and Square with a hole







## Different hole size


```{r}
data <- data2D_pointSquareHole(5000, 0.075)
pl <- myplots(data)
data <- data2D_pointSquareHole(5000, 0.15)
pl <- myplots(data)
data <- data2D_pointSquareHole(5000, 0.3)
pl <- myplots(data)
```

**CONCLUSION 1: We need a distance between barcodes, which doesn't take into account the death time of the hole (the most persistent elements).**


## Different sampling with a hole


```{r}
data <- pointSquareHole(50, 0.3)
pl <- myplots(data)
data <- pointSquareHole(500, 0.3)
pl <- myplots(data)
data <- pointSquareHole(5000, 0.3)
pl <- myplots(data)
```


We need a **distance** between barcodes, which doesn't detect differences bertween the 3 barcodes.

**CONCLUSION 2: less point = less information. The distance has to take the number of points into account**

`pointSquareHole(50, 0.3)` contains a hole but we barely can see it.



## Different sampling when no hole

```{r}
data <- data2D_pointSquareHole(100, 0)
pl <- myplots(data)
data <- data2D_pointSquareHole(300, 0)
pl <- myplots(data)
data <- data2D_pointSquareHole(1000, 0)
pl <- myplots(data)
```

**CONCLUSION 3: we see a kind of limit distribution for H1 elements...**


# Closing or not closing (a circle)





```{r}

data <- data2D_pointCircleGap(100, 0)
pl <- myplots(data)
data <- data2D_pointCircleGap(100, 0.05)
pl <- myplots(data)
data <- data2D_pointCircleGap(100, 0.1)
pl <- myplots(data)
data <- data2D_pointCircleGap(100, 0.3)
pl <- myplots(data)
```


The same with additional Gaussian noise.



```{r}

data<- data2D_pointCircleGap(100, 0)
data$x <- data$x + rnorm(n= 100, mean = 0, sd = 0.1)
data$y <- data$y + rnorm(n = 100, mean = 0, sd = 0.1)
pl <- myplots(data)

data <- data2D_pointCircleGap(100, 0.1)
data$x <- data$x + rnorm(n= 100, mean = 0, sd = 0.1)
data$y <- data$y + rnorm(n = 100, mean = 0, sd = 0.1)
pl <- myplots(data)

data <- data2D_pointCircleGap(100, 0.2)
data$x <- data$x + rnorm(n= 100, mean = 0, sd = 0.1)
data$y <- data$y + rnorm(n = 100, mean = 0, sd = 0.1)
pl <- myplots(data)

data <- data2D_pointCircleGap(100, 0.3)
data$x <- data$x + rnorm(n= 100, mean = 0, sd = 0.1)
data$y <- data$y + rnorm(n = 100, mean = 0, sd = 0.1)
pl <- myplots(data)
```

If no gap, the H1 main element appears at the same time a the last fusion in H0 elements.

The bigger the gap, the bigger the delay in H1.

**CONCLUSION 4: we need a distance that is not sensitive to the gap length between fusion of all H0 and appearence of H1. It has to be only sensitive to the presence/absence of a gap.**

# Comparing different complexes

```{r}

data <- data2D_pointSquareHole(100,0)
pl <- myplots(data, complex = "alpha")
pl <- myplots(data, complex = "rips")
```


# Comparing different noise 

Gaussian versus Uniform noise

```{r}
data<- data2D_pointSquareHole(500, 0)
pl <- myplots(data, complex = "alpha")
data<- data.frame(matrix(0, nrow = 500, ncol = 2))
colnames(data) <- c("x", "y")
data$x <- data$x + rnorm(n= 500, mean = 0, sd = 10)
data$y <- data$y + rnorm(n = 500, mean = 0, sd = 10)
pl <- myplots(data, complex = "alpha")
```

Alpha diagrams (Rips doesn't work...)

```{r ratio law}
data<- data2D_pointSquareHole(10000, 0)
diag1 <- myplots(data)

data<- data.frame(matrix(0, nrow = 10000, ncol = 2))
colnames(data) <- c("x", "y")
data$x <- data$x + rnorm(n= 10000, mean = 0, sd = 10)
data$y <- data$y + rnorm(n = 10000, mean = 0, sd = 10)
diag2 <- myplots(data)
```



```{r}
sel <- diag1[,1] == 1
res1 <- diag1[sel,3]/diag1[sel,2]
sel <- diag2[,1] == 1
res2 <- diag2[sel,3]/diag2[sel,2]
```

Gumbel distribution? 

```{r}
hist(log(log(res1)) - mean(log(log(res1))), breaks = 200)
hist(log(log(res2))- mean(log(log(res1))), breaks = 200)
```


Max values
```{r}
max(res1)
max(res2)
```

The Square with a hole? 

```{r ratio law 2}
data <- data2D_pointSquareHole(10000, 0.3)
diag3 <- myplots(data)
sel <- diag3[,1] == 1
res3 <- diag3[sel,3]/diag3[sel,2]
hist(log(log(res3)) - mean(log(log(res3))), breaks = 200)
```

Max value
```{r}
max(res3)
```



The circle closed + small noise


```{r ratio law 4}
data <- data2D_pointCircleGap(10000, 0)
data$x <- data$x + rnorm(n= 10000, mean = 0, sd = 0.01)
data$y <- data$y + rnorm(n = 10000, mean = 0, sd = 0.01)
diag3 <- myplots(data)
sel <- diag3[,1] == 1
res3 <- diag3[sel,3]/diag3[sel,2]
hist(log(log(res3)) - mean(log(log(res3))), breaks = 200)
```

Max value
```{r}
max(res3)
```



The circle not closed + small noise

```{r ratio law 5}
data <- data2D_pointCircleGap(10000, 0.03)
data$x <- data$x + rnorm(n= 10000, mean = 0, sd = 0.01)
data$y <- data$y + rnorm(n = 10000, mean = 0, sd = 0.01)
diag3 <- myplots(data)
sel <- diag3[,1] == 1
res3 <- diag3[sel,3]/diag3[sel,2]
hist(log(log(res3)) - mean(log(log(res3))), breaks = 200)
```

Max value
```{r}
max(res3)
```



# Conclusions

We need a distance between barcodes, which doesn't take into account the death time of the hole (the most persistent elements)

The distance should be focused on what's happening just after H0 fusions: presence or absence of a gap between main fusions and appearence og H1 elements.

less point = less information. The distance has to take the quantiy of information into account





