---
title: | 
   |  Study of the Sequence of Persistence Diagrams
subtitle:  "Some simple examples"
author: "Vincent Runge"
date: "29/10/2023"
output:
  html_document:
    keep_md: true
    css: styles.css
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
---


```{r setup, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r, message=FALSE}
library(tdaTS) #our package
library(plotly) #3D plotting
library(fields) #image.plot function
```


# data generators

The simple 3D plot function using function `plot_ly` function from the `plotly` package.

```{r}
myplot_3D <- function(data)
{
  minD <- min(data$x, data$y)
  maxD <- max(data$x, data$y)
  M <- max(minD, maxD)
  resScale <- rbind(data,c(0,minD,minD),c(0,minD,maxD),c(0, maxD,minD),c(0,maxD,maxD)) ### add 4 points in the corners for scaling x and y axes
  plot_ly(resScale, x = resScale$x, y = resScale$y, z = resScale$t,
          type="scatter3d", mode = "markers", color = resScale$t)
}
```

We propose, as a starting point, the following four data generators. You give an example for each of the generators. 


## Segment to circle (isometric)


We close a segment into a circle, maintaining the length of the curve (2 pi)

```{r}
data <- segment_to_circle(n = 1000, 
                          change = 0.5, 
                          time_sampling = "unif")

# an example with additional noise

data$x <- data$x + rnorm(n= 1000, mean = 0, sd = 0.2)
data$y <- data$y + rnorm(n = 1000, mean = 0, sd = 0.2)

######## plot the data
myplot_3D(data)

```


## Circle extinction


The circle is smashed into a segment equal to the diameter of the initial circle. 

```{r}
data <- circle_extinction(n = 1000,
                          change = 0.5,
                          time_sampling = "unif")
myplot_3D(data)
```



## Segment to Segments

we split a segment at the changepoint in two at the center point with regular speed with final `gap = 0.1` here.

```{r}
data <- segment_to_segments(n = 1000, 
                            change = 0.5, 
                            gap = 0.1)
myplot_3D(data)
```




## Circle move and distortion

An example with no change. We deform the shape in a ellipse, both rotationg and shifting it. 

```{r}
data <- circle_move_distortion(n = 10000, 
                               rotation = 2, 
                               X_rate = 0.5, 
                               Y_rate = 3)
myplot_3D(data)
```


In most example, there is the possibility to use parameters `X_rate` and/or `Y_rate` to extend or contract the shape and then to change the density of points by level (not their number)

```{r}
data <- segment_to_circle(n = 1000,
                          change = 0.5,
                          X_rate = 0.5,
                          Y_rate = 3)
myplot_3D(data)
```



# Sequence of Persistence Diagrams

In following sections, we will filter the Diagrams using 3 possible thresholds:

```{r}
birth <- 0.15
death <- 0.17
diag <- 0.1
```

The `birth` threshold is a vertical threshold

The `death` threshold is a horizontal threshold

The `diag` threshold is a diagonal threshold

The objective would be to remove the noise part of each of the persistence diagrams.

**We still need to find the right approach to filter the noise (considering filtering by ratio death/birth maybe?)**


## One example segment to Circle

In 3 steps

1. result with no noise

2. result with additional noise

3. segment comparison with matrix of cross Wasserstein distance using dimension 1 elements


```{r, message=FALSE}
set.seed(7)
data <- segment_to_circle(n = 1000, 
                          change = 0.5, 
                          time_sampling = "unif")

Plot_All_Persistence_Diagrams(data,
                              birth = birth,
                              death = death,
                              diagonal = diag,
                              nb_levels = 20)
```



```{r}
data$x <- data$x + rnorm(n= 1000, mean = 0, sd = 0.1)
data$y <- data$y + rnorm(n = 1000, mean = 0, sd = 0.1)

Plot_All_Persistence_Diagrams(data,
                              birth = birth,
                              death = death,
                              diagonal = diag,
                              nb_levels = 20)
```


```{r}
nb_levels <- 20
all_PD <- Generate_All_Persistence_Diagrams(data, 
                                            nb_levels = nb_levels)
################
#### thresholding
################
all_filter <- remove_noisy_points(all_PD,
                                  birth = birth,
                                  death = death,
                                  diagonal = diag,
                                  infinity = FALSE)
################
#### distances: Wasserstein
################
mydim <- 1 #choice for dimension in PD

t(all_filter[,1:2]) %>% as.matrix()

D_wasserstein_2by2 <- distances_Persistence_Diagrams(all_filter, type = "2by2", dimension = mydim)
par(mfrow = c(1,1))

D_wasserstein_2by2 <- t(D_wasserstein_2by2)
image.plot(D_wasserstein_2by2[,nrow(D_wasserstein_2by2):1], axes = F,  col = grey(seq(0,1, length = 256)))
axis(1, at=seq(0,1,length.out = nb_levels), labels= 1:nb_levels)
axis(2, at=seq(0,1,length.out = nb_levels), labels= nb_levels:1)
segments(0,0.5,1,0.5, col = 2, lwd = 4)
segments(0.5,0,0.5,1, col = 2, lwd = 4)
```


## Example circle_extinction


```{r}
set.seed(7)
data <- circle_extinction(n = 2000,
                          change = 0.5,
                          time_sampling = "unif")

Plot_All_Persistence_Diagrams(data,
                              birth = birth,
                              death = death,
                              diagonal = diag,
                              nb_levels = 20)
```



## Example segment_to_segments


```{r}
data <- segment_to_segments(n = 1000, 
                            change = 0.5, 
                            gap = 0.5)

Plot_All_Persistence_Diagrams(data,
                              birth = birth,
                              death = death,
                              diagonal = diag,
                              nb_levels = 20)
```



## Example circle_move_distortion


```{r}
data <- circle_move_distortion(n = 10000, 
                               rotation = 2, 
                               X_rate = 5, 
                               Y_rate = 1, time_sampling = "discrete",
                              nb_levels = 20)

data$x <- data$x + rnorm(n= 10000, mean = 0, sd = 0.1)
data$y <- data$y + rnorm(n = 10000, mean = 0, sd = 0.1)
Plot_All_Persistence_Diagrams(data,
                              birth = birth,
                              death = death,
                              diagonal = diag,
                              nb_levels = 20)


nb_levels <- 20
all_PD <- Generate_All_Persistence_Diagrams(data, 
                                            nb_levels = nb_levels)
################
#### thresholding
################
all_filter <- remove_noisy_points(all_PD,
                                  birth = birth,
                                  death = death,
                                  diagonal = diag,
                                  infinity = FALSE)
################
#### distances: Wasserstein
################
mydim <- 1 #choice for dimension in PD

t(all_filter[,1:2]) %>% as.matrix()

D_wasserstein_2by2 <- distances_Persistence_Diagrams(all_filter, type = "2by2", dimension = mydim)
par(mfrow = c(1,1))

D_wasserstein_2by2 <- t(D_wasserstein_2by2)
image.plot(D_wasserstein_2by2[,nrow(D_wasserstein_2by2):1], axes = F,  col = grey(seq(0,1, length = 256)))
axis(1, at=seq(0,1,length.out = nb_levels), labels= 1:nb_levels)
axis(2, at=seq(0,1,length.out = nb_levels), labels= nb_levels:1)
segments(0,0.5,1,0.5, col = 2, lwd = 4)
segments(0.5,0,0.5,1, col = 2, lwd = 4)
```


# Removing the noise structure 

1. Need to think...

2. Need to look at the litterature 

3. Threshold diagonal ? death over birth ratio ?


## Multiple circles in the 2d plane


```{r}
library(TDA)
circle <- function(n = 1000, center = c(0,0), radius = 1, noise = 0.5)
{
  pos <- runif(n, min = 0, max = 2*pi) 
  x <- center[1] + radius*cos(pos) + rnorm(n, mean = 0, sd = noise)
  y <- center[2] + radius*sin(pos) + rnorm(n, mean = 0, sd = noise)

  res <- matrix(c(x,y), nrow = n, byrow = FALSE)
  return(res)
}

data <- circle()

DiagAlphaCmplx <- alphaComplexDiag(X = data ,
                            library = c("GUDHI", "Dionysus"),
                            location = TRUE,
                            printProgress = FALSE)
plot(DiagAlphaCmplx$diagram, col = 1 + cumsum(DiagAlphaCmplx$diagram[, 1]))


plot(data, col = 1,xaxt="n", yaxt="n",xlab="", ylab="", asp = 1)
    one <- which(DiagAlphaCmplx[["diagram"]][, 1] == 1)
    for (i in seq(along = one))
    {
      for (j in seq_len(dim(DiagAlphaCmplx[["cycleLocation"]][[one[i]]])[1]))
      {
        lines(DiagAlphaCmplx[["cycleLocation"]][[one[i]]][j, , ] + rnorm(n = 4,sd = 0.0), pch = 19, cex = 1, col = i + 1)
      }
    }

dia <-    DiagAlphaCmplx$diagram[DiagAlphaCmplx$diagram[,1] == 1,]

hist(log(dia[,3]/dia[,2]), breaks = 20)
    
```



